Gaussian Processes Autoregressive Models for the Disturbance Storm Time Index

### 2016-05-02 swsc.tex
Updates from ShareLaTeX

For the success of a Gaussian Process model an appropriate choice of kernel function is paramount. The symmetry and positive semi-definiteness of \emph{Gaussian Process} kernels implies that they represent inner-products between some basis function representation of the data. This is analogous to a Fourier series decomposition in Physics. The interested reader is suggested to refer to \cite{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and \citet{hofmann2008} for a thorough treatment on kernel functions and the rich theory behind them. Some common kernel functions used in machine learning are.

\subsubsection{Radial Basis Function (RBF)}

@@

In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.

@@

Before performance bench marks for \emph{OSA} $D_{st}$ prediction are run on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed. For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 and 00:00 15 November 2014 - 23:00 1 December 2014 for training and model selection respectively.

>>>>

For the success of a Gaussian Process model an appropriate choice of kernel function is paramount. The symmetry and positive semi-definiteness of \emph{Gaussian Process} kernels implies that they represent inner-products between some basis function representation of the data. This is analogous to a Fourier series decomposition in Physics. The interested reader is suggested to refer to \cite{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and \citet{hofmann2008} for a thorough treatment on kernel functions and the rich theory behind them. Some common kernel functions used in machine learning include the radial basis function (RBF) kernel $K(\mathbf{x}, \mathbf{y}) = \frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/l^2)$ and the polynomial kernel $K(\mathbf{x}, \mathbf{y}) = (\mathbf{x}^\intercal \mathbf{y} + b)^d$.

The quantities $l$ in the RBF, $b$ and $d$ in the polynomial kernel are known as \emph{hyper-parameters}. Hyper-parameters give flexibility to particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order phenomena respectively. The method of assigning values to the \emph{hyper-parameter} is crucial in the model building process (see section \ref{sec:hyp} below). In this study, we construct Gaussian Process regression models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.

\begin{comment}
\subsubsection{Radial Basis Function (RBF)}

@@

\end{comment}

@@

Before performance bench marks for \emph{OSA} $D_{st}$ prediction are run on the storm events in \cite{Ji2012}, training and model selection of \emph{GP-AR} and \emph{GP-ARX} models on independent data sets must be performed. For this purpose we choose segments 00:00 1 January 2008 - 10:00 11 January 2008 for training and 00:00 15 November 2014 - 23:00 1 December 2014 for model selection.
